---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#imports

library(ggplot2)
library(dplyr)
library(tidyr)
library(colorRamps)
library(reshape2)
library(ggpubr)
library(MASS) 

set.seed(40)
```
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# functions

wright_fisher_selective <- function(p, N, gens, w_11, w_12, w_22) {
  ac = c(p)
  for (x in 1:gens) {
    total_fitness = p * p * w_11 + 2 * p * (1 - p) * w_12 + ((1 - p) ^ 2) * w_22
    p_gf = (p * p * w_11 + p * (1 - p) * w_12)/total_fitness
    p = rbinom(1, 2 * N, p_gf)/(2 * N)
    ac = c(ac, p)
  }
  return(ac)
}

wright_fisher_sim <- function(p, N, gens) {
  answ = c(p)
  for (i in 1:gens){
    p_new = rbinom(1, 2*N, answ[length(answ)]) / (2*N)
    answ = c(answ, p_new)
  } 
  return(answ)
}
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# part 4
#####################################################################
# Case 1) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8
#####################################################################
# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 100
#####################################################################
w_11 = 1
w_12 = 0.9
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y1<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() 

z1 <- ggplot(subset_df, aes(x= value, fill=generation)) + geom_histogram(stat='bin', bins=10,, show.legend = FALSE) + ggtitle('Case 1) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8')

text1<- 'Case 1) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8, n_iter=100, generation = 100'
m1 <- mean(plot_df$value)


#####################################################################
# Case 2) pp fitness = 1, pq fitness = 1, qq fitness = 0.8
#####################################################################

# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 100
#####################################################################
w_11 = 1
w_12 = 1
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y2<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() 

z2 <- ggplot(subset_df, aes(x= value, fill=generation)) + geom_histogram(stat='bin', bins=10,, show.legend = FALSE) + ggtitle('Case 2) pp fitness = 1, pq fitness = 1, qq fitness = 0.8')

text2<- 'Case 2) pp fitness = 1, pq fitness = 1, qq fitness = 0.8, n_iter=100, generations=100'
m2 <- mean(plot_df$value)

#####################################################################
# Case 3) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8
#####################################################################
# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 500
#####################################################################
w_11 = 1
w_12 = 0.9
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y3<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() 

z3 <- ggplot(subset_df, aes(x= value, fill=generation))  + geom_histogram(stat='bin', bins=10,, show.legend = FALSE) + ggtitle('Case 3) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8')

text3<- 'Case 3) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8, n_iter=100, generations=500'
m3 <- mean(plot_df$value)


#####################################################################
# Case 4) pp fitness = 1, pq fitness = 1, qq fitness = 0.8
#####################################################################
# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 500
#####################################################################
# Adding selection to model

w_11 = 1
w_12 = 1
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y4<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() 

z4 <- ggplot(subset_df, aes(x= value, fill=generation))  + geom_histogram(stat='bin', bins=10,, fill='green', show.legend = FALSE) + ggtitle('Case 4) pp fitness = 1, pq fitness = 1, qq fitness = 0.8')

text4<- 'Case 4) pp fitness = 1, pq fitness = 1, qq fitness = 0.8, n_iter=100, generations=500'
m4 <- mean(plot_df$value)


#####################################################################
# Case 5) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8
#####################################################################
# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 1000
#####################################################################
w_11 = 1
w_12 = 0.9
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y5<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line()

z5 <- ggplot(subset_df, aes(x= value, fill=generation))  + geom_histogram(stat='bin', bins=10,, fill='green', show.legend = FALSE) + ggtitle('Case 5) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8')


text5<- 'Case 5) pp fitness = 1, pq fitness = 0.9, qq fitness = 0.8, n_iter=100, generations = 1000'
m5 <- mean(plot_df$value)


#####################################################################
# Case 6) pp fitness = 1, pq fitness = 1, qq fitness = 0.8
#####################################################################
# Adding selection to model
n_iter = 100 # number of iterations
n_gen = 1000
#####################################################################
# Adding selection to model

w_11 = 1
w_12 = 1
w_22 = 0.8
N = 100000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

y6<- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line()

z6 <-ggplot(subset_df, aes(x= value, fill=generation))  + geom_histogram(stat='bin', bins=10,, fill='green', show.legend = FALSE) + ggtitle('Case 6) pp fitness = 1, pq fitness = 1, qq fitness = 0.8')

text6<- 'Case 6) pp fitness = 1, pq fitness = 1, qq fitness = 0.8, n_iter=100, generations = 1000'
m6 <- mean(plot_df$value)# Mean allele frequency in each simulation:


# part 5 code
# Adding selection to model

w_11 = 0.6
w_12 = 1
w_22 = 0.6
N = 1000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

x1 <- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() + 
  scale_y_continuous(limits=c(0, 1))

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

text11<- 'w_11 = 0.6 | w_12 = 1 | w_22 = 0.6 | N = 1000 | p = 0.5 |'
g5 <- mean(plot_df$value)


# Adding selection to model

w_11 = 0.8
w_12 = 1
w_22 = 0.2
N = 1000
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_selective(p, N, n_gen, w_11, w_12, w_22))
  
tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df = melt(tr_df, id.vars='generation')

x2 <- ggplot(plot_df, aes(x=generation, y=value, group=variable)) + geom_line() + 
  scale_y_continuous(limits=c(0, 1))

subset_df <- plot_df %>% filter(generation == nrow(tr_df))

text22<- 'w_11 = 0.8 | w_12 = 1 | w_22 = 0.2 | N = 1000 | p = 0.5'
g6 <- mean(plot_df$value)

```


```{r}
# part 3: wright fisher

jet = matlab.like2(100)

# Simultaing the drift for freq  0.5

n_iter = 250 # number of iterations
n_gen = 20
N = 100 # effective population size
p = 0.5

traj = sapply(1:n_iter, function(x) wright_fisher_sim(p, N, n_gen))

tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df_1 = melt(tr_df, id.vars='generation')

p1 <- ggplot(plot_df_1, aes(x=generation, y=value, group=variable)) + geom_line() + 
  scale_y_continuous(limits=c(0, 1)) + ggtitle('a.) Simultaing the drift for freq  0.5')

# b.) Simultaing the drift for freq  0.1

n_iter = 250 # number of iterations
n_gen = 20
N = 100 # effective population size
p = 0.1

traj = sapply(1:n_iter, function(x) wright_fisher_sim(p, N, n_gen))

tr_df = as.data.frame(traj)
tr_df$generation = 1:nrow(tr_df)

plot_df_2 = melt(tr_df, id.vars='generation')

p2 <- ggplot(plot_df_2, aes(x=generation, y=value, group=variable)) + geom_line() + 
  scale_y_continuous(limits=c(0, 1)) + ggtitle('b.) Simultaing the drift for freq  0.1')



```

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# 2
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```{r}
pp <- 0.775*0.775
qq <- 0.225*0.225
pq_2 <- 2*0.775*0.225
freqs <- c(pp*100,pq_2*100,qq*100)
something <- data.frame("observed" = c(65,25,10), "expected" = as.integer(freqs))
something_again <- data.frame("o_minus_e_squared"=c((something$observed - something$expected)**2))

something_chisq<- sum(something_again/something$expected)

text_for <- 'We see that X-squared is:'
text_for
something_chisq

```


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# 3
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```{r}
condition_zero_1 <- plot_df_1 %>% filter(plot_df_1$value == 0) 
condition_zero_2 <- plot_df_2 %>% filter(plot_df_2$value == 0) 


text <- "Generations where 0's began to be observed:"
text
head(sort(condition_zero_1$generation))
head(sort(condition_zero_2$generation))

```
So in 1st case it never reached 0 while in other we observe 0's as early as 5th generation.

```{r}
ggarrange(p1, p2)
```
We know that, the probability of fixation or loss of an allele in a pure drift Wright-Fisher model only depends on the previous count of the allele and the effective population size.

```{r}
text <- "estimated probability of eliminating an allele that has a starting frequency of a) 0.5; (b) 0.1)"
text
nrow(condition_zero_1)/nrow(plot_df_1)
nrow(condition_zero_2)/nrow(plot_df_2)

```

Intiutively probablities with higher initial frequencies are less likely to be lost for a big enough effective size of populations. As can be seen from the results, when intial p was high (0.5), p0 was 36%. As opposed to when p (0.1), where p0 reached 64%.

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# 4
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

```{r}
ggarrange(z1, z2, z3, z4, z5, z6 + rremove("x.text"),
          ncol = 3, nrow = 2)

```

```

